---
title: "Computational notebooks in R"
output:
  pdf_document: default
---

# Intall the following packages
```{r, eval=FALSE}
install.packages("knitr")
install.packages("rmarkdown")
install.packages("biostatUZH", repos = "http://R-Forge.R-project.org")
```


# Libraries & Variables
```{r, message=FALSE, warning=FALSE}
library(knitr)
library(rmarkdown)
library(biostatUZH)
# PATH_HOME = path.expand("~")
# PATH = file.path(PATH_HOME, "Data", "workshop1")
```

# Load data
```{r, message=FALSE}
# data = read.csv(file = file.path(PATH, "data.csv"), header = TRUE, stringsAsFactors = TRUE)
data = read.csv(file = "data.csv", header = TRUE, stringsAsFactors = TRUE)
```

# 1a) Descriptives of the studies
Table with the number of studies, the median effect size, median sample size, range of publication years, and median number of citations per year per medicall specialty.
```{r}
# number of studies per medical specialty
nrOfStudies = summary(data$specialty)
# median effect size per medical specialty
effect = tapply(data$effect.r, data$specialty, function (x) median(abs(x)))
# median sample size per medical specialty
sampleSize = tapply(data$effect.N, data$specialty, function (x) median(x))
# range of publication years
publicationYears = tapply(data$year, data$specialty, 
                          function (x) sprintf("%d-%d",
                                               range(x, na.rm = TRUE)[1],
                                               range(x, na.rm = TRUE)[2]))
citations = tapply(data$cited.peryear, data$specialty, function (x) median(x, na.rm = TRUE))

# create a data frame
tab.descriptives = data.frame(nrOfStudies, 
                              effect = sprintf("%.2f", effect), 
                              sprintf("%.0f", sampleSize), publicationYears, 
                              citations = sprintf("%.1f", citations)
                              )


kable(tab.descriptives[order(tab.descriptives$nrOfStudies, decreasing = TRUE),])
```

# 1b) Proportion of significant effects per specialty
```{r}
numberSign = tapply(data$effect.p, data$specialty, function (x) sum(x < 0.05))
total = tapply(data$effect.p, data$specialty, length)

tab.effect = data.frame(
  specialty = names(numberSign),
  prop = numberSign/total,
  proportion = sprintf("%.2f",numberSign/total))
rownames(tab.effect) = NULL
kable(tab.effect[,c(1,3)], row.names = FALSE)
```

```{r}
stat = prop.test(numberSign, total, correct = FALSE)
paste("p", formatPval(stat$p.value))
```


# 1c) Logistic model of statistical significance of a study
```{r}
data$specialty = relevel(data$specialty, ref = "Gastroenterology")

fit = glm((effect.p < 0.05) ~ specialty + SJR.Best.Quartile + log2(effect.N) + 
              log2(cited.peryear) + year, data = data, family = "binomial")

tab = tableOR(fit, latex = FALSE, short = TRUE, 
              refLevels = c("Gastroenterology", "Q1"), Wald = TRUE)
kable(tab)
```

