---
title: "Computational notebooks in R"
output:
  pdf_document: default
---

# Intall the following packages
```{r, eval=FALSE}
install.packages("knitr")
install.packages("rmarkdown")
install.packages("biostatUZH", repos = "http://R-Forge.R-project.org")
```


# Libraries & Variables
```{r, message=FALSE, warning=FALSE}
library(knitr)
library(rmarkdown)
library(biostatUZH)
library(Hmisc)
# PATH_HOME = path.expand("~")
# PATH = file.path(PATH_HOME, "Data", "workshop1")
```

# Load data
```{r, message=FALSE}
# data = read.csv(file = file.path(PATH, "data.csv"), header = TRUE, stringsAsFactors = TRUE)
data = read.csv(file = "data.csv", header = TRUE, stringsAsFactors = TRUE)
```

# Analysis 1a: Descriptives of the studies
Table with the number of studies, the median effect size, median sample size, range of publication years, and median number of citations per year per medicall specialty.
```{r}
# number of studies per medical specialty
nrOfStudies = summary(data$specialty)
# median effect size per medical specialty
effect = tapply(data$effect.r, data$specialty, function (x) median(abs(x)))
effect_IRQ_lower = tapply(data$effect.r, data$specialty, function (x) quantile(abs(x), 0.25))
effect_IRQ_upper = tapply(data$effect.r, data$specialty, function (x) quantile(abs(x), 0.75))
# median sample size per medical specialty
sampleSize = tapply(data$effect.N, data$specialty, function (x) median(x))
# range of publication years
publicationYears = tapply(data$year, data$specialty, 
                          function (x) sprintf("%d-%d",
                                               range(x, na.rm = TRUE)[1],
                                               range(x, na.rm = TRUE)[2]))
citations = tapply(data$cited.peryear, data$specialty, function (x) median(x, na.rm = TRUE))

# create a data frame
tab.descriptives = data.frame(nrOfStudies, 
                              effect = sprintf("%.2f (%.2f-%.2f)", effect, effect_IRQ_lower, effect_IRQ_upper), 
                              sprintf("%.0f", sampleSize), publicationYears, 
                              citations = sprintf("%.1f", citations)
                              )

tab = tab.descriptives[order(tab.descriptives$nrOfStudies, decreasing = TRUE),]
colnames(tab) = c("studies", "effect (IRQ)", "N", "publication years", "citations")
kable(tab)
```

# Analysis 1b: Proportion of significant effects per specialty
```{r}
numberSign = tapply(data$effect.p, data$specialty, function (x) sum(x < 0.05))
total = tapply(data$effect.p, data$specialty, length)

myCI =  binconf(x=numberSign, n=total, method = "wilson")

tab.effect = data.frame(
  specialty = names(numberSign),
  prop = numberSign/total,
  lower = myCI[,2],
  upper = myCI[,3],
  proportion = sprintf("%.2f (%.2f-%.2f)",numberSign/total, myCI[,2], myCI[,3]))
rownames(tab.effect) = NULL
kable(tab.effect[order(tab.effect$prop, decreasing = TRUE),c("specialty","proportion")], row.names = FALSE)
```

```{r}
stat = prop.test(numberSign, total, correct = FALSE)
paste("p", formatPval(stat$p.value))
```

```{r}
# simple plot: 
# barplot(tab.effect$prop)
# order levels
idx = order(tab.effect$prop, decreasing = TRUE)
tab.effect$specialty = factor(tab.effect$specialty, levels = tab.effect$specialty[idx])

ggplot(tab.effect, aes(x=specialty, y=prop)) +
  geom_bar(stat="identity", fill="#1b9e77") + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=.5, color="#e7298a") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  ylab("proportion of sign. effects") +
  xlab("medical specialty")
```


# Analysis 1c: Logistic model of statistical significance of a study
```{r}
data$specialty = relevel(data$specialty, ref = "Gastroenterology")

fit = glm((effect.p < 0.05) ~ specialty + SJR.Best.Quartile + log2(effect.N) + 
              log2(cited.peryear) + year, data = data, family = "binomial")

tab = tableOR(fit, latex = FALSE, short = TRUE, 
              refLevels = c("Gastroenterology", "Q1"), Wald = TRUE)
kable(tab)
```

